Sushi monster

### 2025-02-20 inst/sushi.Rmd
sushi

@@

```{r}
a <- as.matrix(read.table("sushi.txt"))
```

```{r,defcalculate_sushi_H}
make_sushi_H <- function(a, strength){
    h <- c(1, strength) # strength of support
    sushi_types <- c("ebi", "anago", "maguro", "ika", "uni", "sake", "tamago", "toro", "tekka", "kappa")
    ec <- # equivalence classes
        c(`ebi`  = 1, `anago` = 2, `maguro` = 1, `ika` = 1, `uni` = 1,
          `sake` = 1, `tamago` = 1, `toro` = 2, `tekka-maki` = 2,
          `kappa-maki` = 1)

    H <- hyper3()

    for(i in seq_len(nrow(a))){
        H <- H + cheering3(sushi_types[a[i,]], e=ec, h=h)
    }
    return(H)
}
```

```{r tryH, cache=TRUE}
wanted <- sample(seq_len(nrow(a)), 10, replace=TRUE, prob=a[,1])
H1888 <- make_sushi_H(a[wanted,-1], strength=1.888)
H1000 <- make_sushi_H(a[wanted,-1], strength=1.000)
```

```{r maxpsushiH,cache=TRUE}
mHa <- maxp(H1888,give=1)
mHb <- maxp(H1000,give=1)
```

```{r showsushiH}
mHa
mHb
```

```{r calclike, cache=TRUE}
f <- function(strength){
   H <- make_sushi_H(a[wanted,-1], strength=strength)
   maxp(H,give=1)$`Log-likelihood`
}
s <- seq(from=0.4,to=2,len=8)
L <- sapply(s,f)
```

```{r}
plot(s,L-max(L),type='b')
```

>>>>

<font size="1"> most of the time is taken calculating the
`calclikesushi` and `maxlike`chunks.  With a sample size of 50 in
chunk `tryH`, and 11 points on the graph, it takes about half an hour
to run without cache. </font>

@@

```{r}
a <- as.matrix(read.table("sushi.txt", skip=22))
set.seed(0)
wanted <- sample(seq_len(nrow(a)),50,replace=TRUE,prob=a[,1])
a <- a[wanted,-1]
```

```{r}
sushi_types <- c("ebi", "anago", "maguro", "ika", "uni", "sake", "tamago", "toro", "tekka", "kappa")
a[] <- sushi_types[a]
noquote(a)
    ec <- c("ebi"  = 1, "anago" = 2, "maguro" = 1, "ika" = 1, "uni" = 1,
          "sake" = 1, "tamago" = 1, "toro" = 2, "tekka" = 2,
          "kappa" = 1)

```

```{r,defcalculate_sushi_H}
make_sushi_H <- function(a, lambda){
    H <- hyper3()

    for(i in seq_len(nrow(a))){
        H <- H + cheering3(a[i,], e=ec, h=c(1,lambda))
    }
    return(H)
}
```

```{r tryH, cache=TRUE}
H1 <- make_sushi_H(a, lambda=1)
H2 <- make_sushi_H(a, lambda=2)
```

```{r maxpsushiH,cache=TRUE}
mH1 <- maxp(H1,give=1)
mH2 <- maxp(H2,give=1)
```

```{r showsushiH}
mH1
mH2
```

```{r calclikesushi, cache=TRUE}
f <- function(lambda){
   H <- make_sushi_H(a, lambda = lambda)
   maxp(H,give=1)$`Log-likelihood`
}

s <- exp(seq(from = log(0.9), to = log(6), len = 11))
L <- sapply(s,f)
```

```{r maxlike, cache=TRUE}
(jj <- optimize(f, 2:3, maximum=TRUE))
```

```{r plotsushilike}
plot(log(s), L-jj$objective, type='b')
```

```{r calc_sushi}
sushi <- make_sushi_H(a, lambda = jj$maximum)
```

```{r calc_sushimaxp}
(sushi_maxp <- maxp(sushi))
```

### Package dataset {-}

Following lines create `sushi.rda`, residing in the `data/` directory of the package.

```{r,label=savesushidataset}
sushi_table <- a
save(sushi_table,  sushi, sushi_maxp, file="sushi.rda")
```
